CF Registry Analysis

### 2023-10-03 Christine Chan/CF Registry/cf_registry_analysis.qmd
updated CFRD registry tables

```{r}
#| include: false
library(knitr)
library(tidyverse)
library(forestmodel)
library(childsds)
```

@@

# Import and clean data
annual <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",

@@

@@

demogs <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")

@@

log_reg <- function(year) {
      A_supplement_feeding, A_feeding2:A_feeding4,
      A_hepatobiliary1_3, contains("A_hepatobiliary2_")

@@

  df$A_feeding2[df$A_supplement_feeding == 0] <- 0
  df$A_feeding3[df$A_supplement_feeding == 0] <- 0
  df$A_feeding4[df$A_supplement_feeding == 0] <- 0
  df$tube_feeding <-
    apply(df[, c("A_feeding2", "A_feeding3", "A_feeding4")], 1, function(r) {

@@

  liver <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
  df$liver_disease <-
    apply(df[, liver], 1, function(r) {

@@

  df$bmi <- df$A_weight / ((df$A_height / 100)^2)
  df$age_perc <- ifelse(df$Age_YrEnd < 20, df$Age_YrEnd, 20)
  df$height_perc <- sds(df$A_height,
    age = df$age_perc,
    sex = df$Gender, male = "M", female = "F",

@@

  df$bmi_perc <- sds(df$bmi,
    age = df$age_perc,
    sex = df$Gender, male = "M", female = "F",

@@

  df$MutClass <- factor(df$MutClass,

@@

  df$NumPulmExacerbation <- cut(df$NumPulmExacerbation,

@@

  df$A_FEV1_pct_predicted <- cut(df$A_FEV1_pct_predicted,

@@

  df$A_FEV1_pct_predicted <- relevel(df$A_FEV1_pct_predicted, ref = "[100,Inf)")
  df$height_perc <- cut(df$height_perc, breaks = c(0, 10, 25, 50, Inf))
  df$height_perc <- relevel(df$height_perc, ref = "(50,Inf]")
  df$bmi_perc <- cut(df$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
  df$bmi_perc <- relevel(df$bmi_perc, ref = "(50,85]")
  df$age_group <- cut(df$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
  df$age_group <- relevel(df$age_group, ref = "[10,18)")

@@

  df[, yn_vars] <- lapply(df[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
  df$A_IsOnEnzymes <- relevel(df$A_IsOnEnzymes, ref = "Yes")

@@

      "A_feeding4", "A_FEV1", liver

@@

@@

>>>>

```{r}
#| include: false
library(Hmisc)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(labelled)
```

@@

# Import data and codebook
annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",

@@

data_dict <- readxl::read_excel("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D

@@

# Get mutations
demogs <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")

@@

# Variable formatting

@@

annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {

@@

liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {

@@

annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",

@@

annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",

@@

annual$MutClass <- factor(annual$MutClass,

@@

annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,

@@

annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,

@@

annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")

@@

annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
log_reg <- function(year, missingness = F) {
      A_supplement_feeding, tube_feeding,
      liver_disease

@@

      "A_feeding4", "A_FEV1", liver_fields

@@

  # Visualize missingness if required
  if (missingness) {
    cat("### Missingness")
    naniar::vis_miss(df)

@@

# Data cleaning

Registry fields included in our liver disease yes/no variable were:

```{r}
data.frame(data_dict$code_meaning[data_dict$var %in% sub("A_", "", liver_fields)]) %>%
  gt() %>%
  tab_options(column_labels.hidden = TRUE)
```

Registry fields included in our tube feeding yes/no variable were:

```{r}
data.frame(data_dict$code_meaning[
  paste0(data_dict$field_name, data_dict$code_value) %in%
    sub("A_", "", tube_fields)
]) %>%
  gt() %>%
  tab_options(column_labels.hidden = TRUE)
```

# CFRD validity check

```{r}
annual %>%
  mutate(cfrd = as.numeric(cfrd_status_annual == "CFRD+")) %>%
  group_by(eDWID) %>%
  summarise(
    `CFRD Status` = any(rle(cfrd)$lengths[rle(cfrd)$values == 1] >= 2),
    .groups = "drop"
  ) %>%
  mutate(`CFRD Status` = factor(`CFRD Status`,
    levels = c(F, T),
    labels = c("CFRD-", "CFRD+")
  )) %>%
  count(`CFRD Status`) %>%
  gt()
```

# HbA1c by year

```{r}
annual %>%
  group_by(ReviewYear) %>%
  summarise(
    `Num. HbA1c Tests` = sum(!is.na(A_hgba1c)),
    `Total n` = n(),
    `Percent with HbA1c` = round(sum(!is.na(A_hgba1c)) / n() * 100, 2)
  ) %>%
  gt()
```

## 2004

```{r results='asis',fig.height=10,fig.width=10}
log_reg(2004,missingness = T)
```
